### 2016-03-14 paper/icfp16-submission/evaluation.tex
move "jump compression" to interactive-short
move lem:context-compat to the appendix

\begin{figure}[t]
\centering
\begin{minipage}{\linewidth}
\includegraphics[width=\linewidth]{coverage.png}
\end{minipage}
\begin{minipage}{\linewidth}
\includegraphics[width=\linewidth]{distrib.png}
\end{minipage}
% \vspace{-8ex}
\caption{Results of our coverage testing and the distribution of test
  outcomes. Our random search successfully finds witnesses for 83\% of
  the programs in under one second, improving slightly to 87\% in under
  10 seconds. In both datasets we detect actual type errors about 82\%
  of the time, unbound variables or constructors 3-4\% of the time, and
  diverging loops 2-3\% of the time. For the remaining 11-12\% of the
  programs we are unable to provide any useful feedback.  }
\label{fig:results-witness}
\end{figure}

@@

\paragraph{Jump Compression}
%
A \emph{jump compressed} trace is one whose edges are limited to forward
or backward jumps. (The sizes of these traces is equal to the jump
metric below.)
%
In our experience, jump compression abstracts many details of the
computation that are often uninteresting or irrelevant to the
explanation.
% 
In particular, jump compressed traces hide low-level operations and
summarize function calls as call-return pairs, see
Figure~\ref{fig:fac-jump} for a variant of @fac@ that implements the
subtraction as a function call instead of a primitive.
%
\begin{figure}[t]
\centering
\includegraphics[height=2in]{fac-minus.png}
\caption{Jump-compressed trace of \texttt{fac 1} with subtraction
  implemented as a function call.}
\label{fig:fac-jump}
\end{figure}
%
Once the user has identified an interesting call-return pair, they can
step into the call and proceed with more fine-grained steps.
%
% Expanding the initial visualization to a jump-compressed trace is often
% an ideal first step;
% % expanding a jump-compressed traces ideal for the initial state of the
% they provide a concise overview of the execution and allow the user to
% step directly into an interesting function call, at which point the can
% proceed with more fine-grained steps.
%
Note that jump compressed traces are not quite the same as
stack-traces as they show \emph{all} function calls, including
those that returned successfully.
%
For example, the display on the right in Figure~\ref{fig:factorial} shows
a jump-compressed trace, of size 4. The full trace for
the same example has size 20.

\paragraph{Metrics} Thus, our two metrics are:
% size of the full trace,
% \ie the number of small-step reductions, and the size of the jump-compressed
% version of the trace.
%
\begin{enumerate}
\item \emphbf{Single-Step Metric} The size of the trace after expanding
  all of the single-step edges from the witness to the stuck term, and
  % This can be thought of as a worst-case
  % complexity, \ie ``How big is the fully-expanded trace?''
\item \emphbf{Jump Metric} The size of the trace for after expanding
  only the forward (or backward) jumps from the witness to the stuck
  term.
\end{enumerate}

>>>>

\begin{figure*}[t]
\centering
\begin{minipage}{\linewidth}
\centering
\includegraphics[width=0.49\linewidth]{coverage.png}
% \end{minipage}
% \begin{minipage}{\linewidth}
\includegraphics[width=0.49\linewidth]{distrib.png}
\end{minipage}
% \vspace{-8ex}
\caption{Results of our coverage testing and the distribution of test
  outcomes. Our random search successfully finds witnesses for 83\% of
  the programs in under one second, improving slightly to 87\% in under
  10 seconds. In both datasets we detect actual type errors about 82\%
  of the time, unbound variables or constructors 3-4\% of the time, and
  diverging loops 2-3\% of the time. For the remaining 11-12\% of the
  programs we are unable to provide any useful feedback.  }
\label{fig:results-witness}
\end{figure*}

@@

% \paragraph{Metrics} Thus, our two metrics are:
% size of the full trace,
% \ie the number of small-step reductions, and the size of the jump-compressed
% version of the trace.
%
\begin{enumerate}
\item \emphbf{Single-step Metric} The size of the trace after expanding
  all of the single-step edges from the witness to the stuck term, and
  % This can be thought of as a worst-case
  % complexity, \ie ``How big is the fully-expanded trace?''
\item \emphbf{Jump-compressed Metric} The size of the jump-compressed trace.
\end{enumerate}

### 2016-03-14 paper/icfp16-submission/gallery2.tex
fixed clobbering
move "jump compression" to interactive-short
move lem:context-compat to the appendix

Next, we present a \emph{qualitative} evaluation that compares
the explanations provided by \toolname's dynamic witnesses with
the static reports produced by the \ocaml\ compiler and \sherrloc,
a state-of-the-art fault localization approach~\cite{zhang_toward_2014}.
%
In particular, we illustrate, using a series of examples drawn
from student programs in the \ucsdbench\ dataset how \toolname's
jump-compressed traces can get to the heart of the error by
%
highlighting the conflicting values that cause the program to get
stuck, rather that blaming a single one,
%
showing the steps necessary to reach the stuck state, and
%
not assuming that a function is correct just because it type-checks.
%
For each example we will present
(1)~the code,
(2)~the error message returned \ocaml,
(3)~the error locations returned by \ocaml\ (underlined) and \sherrloc\ (in bold),
\footnote{When the locations from \ocaml\ and \sherrloc\ overlap,
we just underline the relevant code.}
(4)~the jump-compressed trace produced by \toolname.

@@

\paragraph{Example: Recursion with Bad Base Case}
%
The function @sumList@ should add up
the elements of its input list.
%
\begin{ecode}
  let rec sumList xs = match xs with
    | []    -> ==[]==
    | y::ys -> y + __sumList ys__
\end{ecode}
%
Unfortunately, it incorrectly returns @[]@
in the base case instead of @0@.
%
\sherrloc\ blames the base case, and \ocaml\
assumes the base case is correct and blames
the \emph{recursive call} on line 3:
%
\begin{verbatim}
  This expression has type
    'a list
  but an expression was expected of type
    int
\end{verbatim}
%
Both of the above are parts of the full story, which
is succinctly summarized by \toolname's trace showing
how @sumList [1; 2]@ gets stuck at @2 + []@.
%
\begin{center}
  \includegraphics[height=125px]{sumlist.png}
\end{center}
%
The trace clarifies immediately (via the third step)
that the @[]@ is the result of the recursive call
@sumList []@, and shows how it is incompatible with
the subsequent \texttt{+} operation.

>>>>

Next, we present a \emph{qualitative} evaluation that compares
the explanations provided by \toolname's dynamic witnesses with
the static reports produced by the \ocaml\ compiler and \sherrloc,
a state-of-the-art fault localization approach~\cite{zhang_toward_2014}.
%
In particular, we illustrate, using a series of examples drawn
from student programs in the \ucsdbench\ dataset how \toolname's
jump-compressed traces can get to the heart of the error by
%
highlighting the conflicting values that cause the program to get
stuck, rather that blaming a single one,
%
showing the steps necessary to reach the stuck state, and
%
not assuming that a function is correct just because it type-checks.
%
For each example we will present
(1)~the code,
(2)~the error message returned \ocaml,
(3)~the error locations returned by \ocaml\ (underlined) and \sherrloc\ (in bold)
\footnote{When the locations from \ocaml\ and \sherrloc\ overlap,
we just underline the relevant code.},
(4)~the jump-compressed trace produced by \toolname.

@@

\paragraph{Example: Recursion with Bad Base Case}
%
The function @sumList@ should add up
the elements of its input list.
%
\begin{ecode}
  let rec sumList xs = match xs with
    | []    -> ==[]==
    | y::ys -> y + __sumList ys__
\end{ecode}
%
Unfortunately, in the base case, it returns @[]@
instead of @0@.
%
\sherrloc\ blames the base case, and \ocaml\
assumes the base case is correct and blames
the \emph{recursive call} on line 3:
%
\begin{verbatim}
  This expression has type
    'a list
  but an expression was expected of type
    int
\end{verbatim}
%
Both of the above are parts of the full story, which
is succinctly summarized by \toolname's trace showing
how @sumList [1; 2]@ gets stuck at @2 + []@.
%
\begin{center}
  \includegraphics[height=125px]{sumlist.png}
\end{center}
%
The trace clarifies immediately (via the third step)
that the @[]@ is the result of the recursive call
@sumList []@, and shows how it is incompatible with
the subsequent \texttt{+} operation.

### 2016-03-14 paper/icfp16-submission/interactive-short.tex
fixed genWitness
move "jump compression" to interactive-short
move lem:context-compat to the appendix

>>>>

\paragraph{Jump Compression}
%
A \emph{jump compressed} trace is one whose edges are limited to forward
or backward jumps.
% (The sizes of these traces is equal to the jump
% metric below.)
%
In our experience, jump compression abstracts many details of the
computation that are often uninteresting or irrelevant to the
explanation.
%
In particular, jump compressed traces hide low-level operations and
summarize function calls as call-return pairs, see
Figure~\ref{fig:fac-jump} for a variant of @fac@ that implements the
subtraction as a function call instead of a primitive.
%
\begin{figure}[t]
\centering
\includegraphics[height=2in]{fac-minus.png}
\caption{Jump-compressed trace of \texttt{fac 1} with subtraction
  implemented as a function call.}
\label{fig:fac-jump}
\end{figure}
%
Once the user has identified an interesting call-return pair, they can
step into the call and proceed with more fine-grained steps.
%
% Expanding the initial visualization to a jump-compressed trace is often
% an ideal first step;
% % expanding a jump-compressed traces ideal for the initial state of the
% they provide a concise overview of the execution and allow the user to
% step directly into an interesting function call, at which point the can
% proceed with more fine-grained steps.
%
Note that jump compressed traces are not quite the same as
stack-traces as they show \emph{all} function calls, including
those that returned successfully.
%
For example, the display on the bottom-left in Figure~\ref{fig:factorial} shows
a jump-compressed trace, of size 4. The full trace for
the same example (on the right of the Figure~\ref{fig:factorial}) has size 19.

% \RJ{add para about jump-compressed traces + example. include full trace and compressed trace.
% could merge with example in overview.}

### 2016-03-14 paper/icfp16-submission/overview.tex
aasdasd
move "jump compression" to interactive-short
move lem:context-compat to the appendix

To defer input generation, we borrow the notion of a ``hole'' from
SmallCheck~\cite{runciman_smallcheck_2008}.
%
A hole --- written \vhole{\thole} --- is a \emph{placeholder} for a
value of some unknown type \thole.
%
We leave all inputs as uninstantiated holes until they are demanded by
the program, \eg due to a primitive operation like the @<=@ test.

@@

\subsection{Visualizing Witnesses}
\label{sec:visual-witness}
We have shown how to reliably find witnesses to type errors in \ocaml,
but this does not fully address our original goal --- to \emph{explain}
the errors.
%
Having identified an input vector that triggers a bug, a common next
step is to step through the program with a \emph{debugger} to observe
how the program evolves.
%
The existing debuggers and interpreters for \ocaml\ assume a type-correct
program, so unfortunately we cannot use them off-the-shelf.
%
Instead we extend our search for witnesses to produce an execution
trace.

@@

>>>>

To defer input generation, we borrow the notion of a ``hole'' from
SmallCheck~\cite{runciman_smallcheck_2008}.
%
A hole --- written \vhole{\thole} --- is a \emph{placeholder} for a
value \ehole of some unknown type \thole.
%
We leave all inputs as uninstantiated holes until they are demanded by
the program, \eg due to a primitive operation like the @<=@ test.

@@

\subsection{Visualizing Witnesses}
\label{sec:visual-witness}
We have shown how to reliably find witnesses to type errors in \ocaml,
but this does not fully address our original goal --- to \emph{explain}
the errors.
%
Having identified an input vector that triggers a crash, a common next
step is to step through the program with a \emph{debugger} to observe
how the program evolves.
%
The existing debuggers and interpreters for \ocaml\ assume a type-correct
program, so unfortunately we cannot use them off-the-shelf.
%
Instead we extend our search for witnesses to produce an execution
trace.

@@

\paragraph{Jump-compressed Witnesses}
It is rare for the initial state of the visualization to be
informative enough to diagnose the error.
%
We provide a short-cut to expand the witness to a \emph{jump-compressed}
trace, which contains every function call and return step.
%
The jump-compressed trace abstracts the computation as a sequence of
call-response pairs, providing a high-level overview of steps taken
to reach the crash, and a high level of compression compared to the
full trace.
%
For example, the jump-compressed trace in Figure~\ref{fig:factorial}
contains 4 nodes compared to the 19 in the fully expanded trace.
%
Our benchmark suite of student programs shows that jump-compression is
practical, with an average trace size of 7 nodes and a median of 5.

### 2016-03-14 paper/icfp16-submission/semantics.tex
move "jump compression" to interactive-short
move lem:context-compat to the appendix

Next, we formalize the notion of type error witnesses as follows.
%
First, we define a core calculus within which we will work~(\S~\ref{sec:syntax}).
%
Second, we develop a (non-deterministic) operational semantics
for ill-typed programs that precisely defines the notion
of a \emph{witness}~(\S~\ref{sec:semantics}).
%
Third, we formalize and prove a notion of \emph{generality} for
witnesses, which states, intuitively, that if we find a
single witness then for \emph{every possible} type
assignment there exist inputs that are guaranteed to make
the program ``go wrong''~(\S~\ref{sec:soundness}).
%
Finally, we refine the operational semantics into a
\emph{search procedure} that returns concrete (general)
witnesses for ill-typed programs~\S~(\ref{sec:search-algorithm}).

@@

\ES{Where does this fit in??}
Next, we introduce a notion of type % and value
contexts, defined analogously to evaluation contexts.
%
$$
\begin{array}{lcl}
  T &::=& \bullet \spmid \ttree{T} \spmid \tprod{T}{t} \spmid \tprod{t}{T} \\
  % V &::=& \bullet \spmid \vnode{t}{V}{v}{v} \spmid \vnode{t}{v}{V}{v} \spmid \vnode{t}{v}{v}{V} \\
\end{array}
$$
%

\begin{lem}
\label{lem:context-compat}
For all types $s$ and $t$,
if $\tcompat{\intctx{s}}{t}$
then $t = \intctx{s'}$
such that $\tcompat{s}{s'}$.
\end{lem}
\begin{proof}
  By induction on $t$.
\end{proof}

@@

>>>>

Next, we formalize the notion of type error witnesses as follows.
%
First, we define a core calculus within which we will work~(\S~\ref{sec:syntax}).
%
Second, we develop a (non-deterministic) operational semantics
for ill-typed programs that precisely defines the notion
of a \emph{witness}~(\S~\ref{sec:semantics}).
%
Third, we formalize and prove a notion of \emph{generality} for
witnesses, which states, intuitively, that if we find a
single witness then for \emph{every possible} type
assignment there exist inputs that are guaranteed to make
the program ``go wrong''~(\S~\ref{sec:soundness}).
%
Finally, we refine the operational semantics into a
\emph{search procedure} that returns concrete (general)
witnesses for ill-typed programs~\S~(\ref{sec:search-algorithm}).
%
We have formalized and tested our semantics and generality theorem
in \textsc{PLT-Redex}~\cite{felleisen2009semantics}.

@@

@@

  % Next, we introduce a notion of type % and value
  % contexts, defined analogously to evaluation contexts.
  % %
  % $$
  % \begin{array}{lcl}
  %   T &::=& \bullet \spmid \ttree{T} \spmid \tprod{T}{t} \spmid \tprod{t}{T} \\
  %   % V &::=& \bullet \spmid \vnode{t}{V}{v}{v} \spmid \vnode{t}{v}{V}{v} \spmid \vnode{t}{v}{v}{V} \\
  % \end{array}
  % $$
